---
title: "Demographics and reproductive biology of H. schistosus from bycatch on the west coast of India"
subtitle: "Analysis for Manuscript"
author: "Shawn Dsouza and Chetan Rao" 
output:
  rmdformats::robobook:
    code_folding: hide
    css: style.css
    df_print: kable
---

```{r setup, include=FALSE}
#Global options

options(max.print="75")
knitr::opts_chunk$set(error = T, warning = F, message = F, tidy = T, cache = T,
                      dpi = 300, fig.align = "center", fig.height = 4.5, fig.width = 8, digits = 3)

require(ggplot2)

theme_set(theme_bw()+
            theme(legend.position = "top",
                  text = element_text(size = 15)
                  )
          )#setting ggplot theme

```

# Introduction

This the code for analysis and supplementary figures for the manuscript. The manuscript is currently under review.

```{r}
#Required libraries

require(tidyverse)
require(lubridate)
require(viridis)

# Importing data

snakes = read.csv("./Data/Sea-snakes_fish-dep_mastersheet_250420.csv")
embryos = read.csv("./Data/Sea snake_embryo data_mastersheet.csv")

snakes <- snakes%>%
  mutate(Year = as.factor(Year))%>%
  filter(Species == "Hydrophis curtus" | Species == "Hydrophis schistosus")%>%
  droplevels()

HS <- snakes%>%
  filter(Species == "Hydrophis schistosus")%>% # filtering out other snakes from bycatch data
  mutate(Year = as.factor(Year))

# function to calculate mcfadden's r2

mcfadden <- function(x){
  
  r2 <- 1 - (x$deviance/x$null.deviance)
  
}

# sampling years
years <- data.frame(Year = factor(c(2016:2019)))
```

## Sample size

The number of snakes (N) sampled from bycatch of trawlers, shore seines and gillnetss.

```{r}
snakes%>%
  group_by(Species)%>%
  count(name = "N")
```

The number of snakes sampled in each year:

```{r}
n.yr <- snakes%>%
  group_by(Species, Year)%>%
  count(name = "N")

n.yr%>%
  spread(Year, N)
```

The number of trips (n) and fishing effort (haul-hours) sampled for sea snakes in the current study.

```{r}
snakes%>%
  filter(Gear.Type != "")%>%
  select(Date, Boat.ID, Gear.Type, No..of.Hauls, Average.Haul.Duration..Hours., Tow.duration.hours)%>%
  distinct()%>%
  # Calculating tow duration
  mutate(Tow.duration.hours = case_when(!is.na(Tow.duration.hours) ~ Tow.duration.hours,
                                        is.na(Tow.duration.hours) ~ No..of.Hauls*Average.Haul.Duration..Hours.))%>%
  group_by(Gear.Type)%>%
  summarise(n = n(), # counting number of trips sampled
            haul.hours = sum(Tow.duration.hours, na.rm = T)) # total fishing effort
  
```

# Demographics of *H. schistosus* across time

## Age structure

```{r}
# calculating the mean SVL across years

age.yr <- snakes%>%
  group_by(Species, Year)%>%
  summarise(N = n(),
            mean = mean(Snout.to.Vent..cm., na.rm = T))

# SVL cut off for age classes

maturtity <- snakes%>%
  group_by(Species)%>%
  count()%>%
  mutate(juv = 35,
         adult = ifelse(Species == "Hydrophis curtus", 54, 65))
  
```

### Distribution fo SVL across years

```{r fig.height=9}
# plotting distribution of SVL across years

snakes%>%
  filter(Snout.to.Vent..cm. > 20)%>%
  ggplot(aes(Year, Snout.to.Vent..cm.))+
  geom_violin(fill = "grey")+
  geom_boxplot(width = 0.1)+
  geom_hline(data = maturtity, aes(yintercept = adult), linetype = "dashed")+
  geom_hline(data = maturtity, aes(yintercept = juv), linetype = "dotted")+
  #stat_summary(fun.data = "mean_sdl", geom = "pointrange", size = 1)+
  geom_label(data = age.yr, aes(Year, 10, label = N))+
  facet_wrap(~Species, ncol = 1, scales = "free_y")+
  labs(y = "Snout to vent length (cm)")+
  theme(strip.text = element_text(face = "italic"))

ggsave(last_plot(), filename = "./Figures/figure1.tiff", height = 9, width = 8)
  
```

We did not find any *H. curtus* neonates in our sampling. Majority of *H. curtus* were juveniles. *H. schistosus* populations consisted of mostly adults.

Linear model for change in SVL of each species across years:

```{r}

library(car)
require(broom)

snakes%>%
  group_by(Species)%>%
  select(Year, Snout.to.Vent..cm.)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(Snout.to.Vent..cm. ~ Year, data = .)),
          sumr = map(mod, broom::tidy),
         stat = map(mod, glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(c(Species:r.squared))
```

ANOVA to test for effect of years:

```{r}
snakes%>%
  group_by(Species)%>%
  select(Year, Snout.to.Vent..cm.)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(Snout.to.Vent..cm. ~ Year, data = .)),
          stat = map(mod, car::Anova))%>%
  select(stat)%>%
  unnest()
```

Plotting fitted values:

```{r}

snakes%>%
  group_by(Species)%>%
  select(Year, Snout.to.Vent..cm.)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(Snout.to.Vent..cm. ~ Year, data = .)),
         fit = map(mod, ~predict(., years, se.fit = T)),
         fit = map(fit, ~as.data.frame(.)))%>%
  select(fit)%>%
  unnest()%>%
  mutate(Year = factor(c(2016:2019)))%>%
  ggplot(aes(Year, fit))+
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit))+
  labs(y = "SVL (cm)")+
  facet_wrap(~Species)+
  theme(strip.text = element_text(face = "italic"))
```

There was a significant change in SVL distribution of species across years.

### Proportion of developmental classes across years

```{r}
snakes%>%
  filter(!is.na(Class))%>%
  group_by(Species)%>%
  # Counting number of individuals in each developmental class
  count(Class)%>%
  complete(Class, fill = list(n = 0))%>%
  # Total number sampled in each year
  mutate(N = sum(n))%>%
  group_by(Species, Class)%>%
  # Proportion of each developmental class
  mutate(p = n/N,
         sd = sqrt(p*(1-p)/N))

```

### Testing change in proportion of developmental classes across years

Fitting a multinomial logistic model to determine change in age classes across years in each species.

```{r results=F}
require(nnet)
require(broom)

dev.yr <- snakes%>%
  select(Species, Year, Class)%>%
  filter(!is.na(Class))%>% # removing empty class 
  group_by(Species)%>%
  nest()%>%
  mutate(mod = map(data, ~multinom(Class ~ Year, data = .)),
         null = map(data, ~multinom(Class ~ 1, data = .)),
         smry = map(mod, ~broom::tidy(., exonentiate = T)),
         # Calculating mc fadden's r2
         r2 = map2_dbl(mod, null, ~ 1 - .x$value/.y$value))
```

```{r}
dev.yr%>%
  select(smry, r2)%>%
  unnest()
```

ANOVA to test for effect of years:

```{r}
dev.yr%>%
  mutate(stat = map(mod, car::Anova))%>%
  select(stat)%>%
  unnest()
```

Plotting fitted estimates of proportions:

```{r}
dev.yr%>%
  mutate(fit = map(mod, ~data.frame(predict(., years, "probs"))))%>%
  select(fit)%>%
  unnest()%>%
  mutate(Year = factor(c(2016:2019)))%>% # adding years variable
  left_join(n.yr, by = c("Species", "Year"))%>% # getting sample size in each year
  gather(c(Adult, Juvenile, Neonate), key = "Class", value = "p.hat")%>%
  mutate(se = sqrt(p.hat*(1-p.hat)/N))%>% # caluclating standard error
  # Plotting
  ggplot(aes(Year, p.hat, color = Species))+
  geom_pointrange(aes(ymin = p.hat-se, ymax = p.hat+se))+
  labs(y = "Estimated proportion")+
  facet_wrap(~Class)+
  theme(legend.text = element_text(face = "italic"))
```

Age structure of *H. schistosus* does not change significantly over a four year period from 2016 to 2018.

The proportion of *H. curtus* juveniles increases significantly between 2016 and 2018

### Change in SVL distribution over months

```{r, fig.height = 4.5, fig.width=8}
month.svl <- HS%>%
  filter(Year == "2018", # data not sufficient for other years
         Snout.to.Vent..cm. > 20)%>% # removing erroneous data
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(Month)%>%
  group_by(Month)%>%
  # Calculating mean SVL in each month
  summarise(mean.SVL = mean(Snout.to.Vent..cm., na.rm = T))

# Month of observed births

births <- data.frame(Species = "Hydrophis schistosus", Month = "April")
```

```{r, fig.height = 4.5, fig.width=8}
# plotting distribution of SVL across months

HS%>%
  filter(Year == "2018", # data not sufficient for other years
         Snout.to.Vent..cm. > 20)%>% # removing erroneous data
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(nesting(Species), Month)%>%
  droplevels()%>%
  ggplot(aes(Month, Snout.to.Vent..cm.))+
  geom_violin(fill = "light grey")+
  #geom_point(data = month.svl, aes(x = Month, y = mean.SVL), size = 3)+
  geom_boxplot(width = 0.1)+
  geom_segment(data = births, aes(x = Month, xend = Month, y = 0 , yend = 10), #marking births
               arrow = arrow(length = unit(0.25, "cm"), ends = "first"), size = 1)+
  geom_text(data = births, aes(x = Month, y = 20, label = paste("Observed \n births")))+
  geom_vline(aes(xintercept = "June"), size = 1)+#start of the monsoon ban
  geom_vline(aes(xintercept = "August"), size = 1)+#end of the monsoon ban
  geom_text(aes(x = "July", y = 80, label = "Monsoon ban"))+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(y = "Snout to vent length (cm)")

```

Proportion of neonates increases in March to May. Birth and neonates were observed around the same time.

### Proportion of neonates in each month of 2018

```{r}
snakes%>%
  filter(Year == "2018", # data not sufficient for other years
         Snout.to.Vent..cm. > 20)%>% # removing erroneous data
  mutate(Month = factor(Month, levels = month.name))%>%
  group_by(Species, Month)%>%
  summarise(p = sum(Snout.to.Vent..cm. < 40)/n(),
            sd = sqrt(p*(1-p)/n()))
```

## Sex ratios

### Proportion of females encountered by species

```{r}
snakes%>%
  group_by(Species)%>%
  filter(Sex == "Male" | Sex == "Female")%>%
  summarise(females = sum(Sex == "Female"),
            N = n(),
            P = females/N,
            sd = sqrt(P*(1-P)/N))
```

### Testing equality in sex ratios

```{r}
# is proportion of females different from 0.5?

snakes%>%
  filter(Sex == "Male" | Sex == "Female")%>%
  group_by(Species)%>%
  summarise(females = sum(Sex == "Female"),
            N = n())%>%
  group_by(Species)%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$females, .$N, p = 0.5)), # Z - test for proportion
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()%>%
  select(-c(method, alternative))
```

Sex ratio is not different from 0.5; p = 0.71.

### Testing shifts in sex ratios over years

```{r }
snakes%>%
  filter(Sex == "Male" | Sex == "Female")%>%
  group_by( Species, Year)%>%
  summarise(N = n(),
            females = sum(Sex == "Female"),
            prop.female = females/N, 
            sd = sqrt(prop.female*(1-prop.female)/N))%>%
  ggplot(aes(Year, prop.female, fill = Species))+
  geom_col(width = 0.5, col = "black", position = position_dodge())+
  geom_errorbar(aes(ymin = prop.female-sd, ymax = prop.female + sd), position = position_dodge(width = 0.5), width = 0.25)

```

Logistic model to test for change in sex ratios of species across years:

```{r}
snakes%>%
  select(Species, Year, Sex)%>%
  filter(Sex == "Male" | Sex == "Female")%>% # removing empty sex
  mutate(Sex = factor(Sex, levels = c("Male", "Female")))%>%
  group_by(Species)%>%
  nest()%>%
  mutate(mod = map(data, ~glm(Sex ~ Year, family = binomial, data = .)),
         sumry = map(mod, ~tidy(.)),
         r2 = map_dbl(mod, ~mcfadden(.)))%>%
  select(sumry, r2)%>%
  unnest(sumry)
```

Plotting predicted values:

```{r}
snakes%>%
  select(Species, Year, Sex)%>%
  filter(Sex == "Male" | Sex == "Female")%>% # removing empty sex
  mutate(Sex = factor(Sex, levels = c("Male", "Female")))%>%
  group_by(Species)%>%
  nest()%>%
  mutate(mod = map(data, ~glm(Sex ~ Year, family = "binomial", data = .)),
         fit = map(mod, ~predict(., years, se.fit = T, type = "response")),
         fit = map(fit, ~data.frame(.)))%>%
  select(fit)%>%
  unnest(fit)%>%
  mutate(Year = factor(c(2016:2019)))%>%
  ggplot(aes(Year, fit))+
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit))+
  labs(y = "P(Female|Year)")+
  facet_wrap(~Species)+
  theme(strip.text = element_text(face = "italic"))
```

The proportion of female *H. curtus* changed significantly across years. *H. schistosus* sex ratios remained constant across sampled years.

# Mortality in bycatch

## Overall mortality rate

```{r}
snakes%>%
  group_by(Species)%>%
  count(Condition.at.encounter..D.A.)%>%
  # Calculating mortality rate
  mutate(N = sum(n),
         prop.dead = n/N,
         sd = sqrt(prop.dead*(1-prop.dead)/N))%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

### Testing difference in mortality across species

Logistic model:

```{r}
snakes%>%
  filter(Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Species, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")))%>%
  nest()%>%
  mutate(test = map(data, ~glm(Condition ~ Species, family = "binomial", data = .)),
         sumr = map(test, ~broom::tidy(.)),
         stat = map(test, broom::glance),
         r2 = map(test, ~mcfadden(.)))%>%
  select(sumr, stat, r2)%>%
  unnest()%>%
  select(term:p.value, df.residual, r2)
```

Fitted probability of mortality:

```{r}

spp = data.frame(Species = c("Hydrophis schistosus", "Hydrophis curtus"))

snakes%>%
  filter(Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Species, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")))%>%
  nest()%>%
  mutate(mod = map(data, ~glm(Condition ~ Species, family = "binomial", data = .)),
         fit = map(mod, ~predict(., spp, se.fit = T, type = "response")),
         fit = map(fit, ~data.frame(.)))%>%
  select(fit)%>%
  unnest(fit)%>%
  mutate(Species = factor(c("Hydrophis schsitosus", "Hydrophis curtus")))%>%
  ggplot(aes(Species, fit))+
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit))+
  labs(y = "P(Dead|Species)")+
  theme(axis.text.x =  element_text(face = "italic"))

```

*H. curtus* had as significatly higher mortality rate in bycatch than *H. schistosus*

## Mortality rate by age class

Are juveniles more vulnerable to bycatch mortality than adults?

```{r}

snakes%>%
  filter(!is.na(Class))%>%
  group_by(Species, Class)%>%
  count(Condition.at.encounter..D.A.)%>%
  #calculating mortality rate
  mutate(N = sum(n),
         prop.dead = n/N,
         sd = sqrt(prop.dead*(1-prop.dead)/N))%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

### Testing difference in mortality rates across age classes within species

Logistic model:

```{r}

snakes%>%
  filter(!is.na(Class),
         Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Species, Class, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")))%>%
  group_by(Species)%>%
  nest()%>%
  mutate(test = map(data, ~glm(Condition ~ Class, family = "binomial", data = .)),
         sumr = map(test, ~broom::tidy(.)),
         stat = map(test, broom::glance),
         r2 = map(test, ~mcfadden(.)))%>%
  select(sumr, stat, r2)%>%
  unnest()%>%
  select(term:p.value, df.residual, r2)

```

Fitted probability of mortality across age classes:

```{r}

dev = data.frame(Class = c("Adult", "Juvenile", "Neonate"))

snakes%>%
  filter(!is.na(Class),
         Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Species, Class, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")))%>%
  group_by(Species)%>%
  nest()%>%
  mutate(mod = map(data, ~glm(Condition ~ Class, family = "binomial", data = .)),
         fit = map(mod, ~predict(., dev, se.fit = T, type = "response")),
         fit = map(fit, ~data.frame(.)))%>%
  select(fit)%>%
  unnest(fit)%>%
  mutate(Class = factor(c("Adult", "Juvenile", "Neonate")))%>%
  ggplot(aes(Class, fit))+
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit))+
  labs(y = "P(Dead)")+
  facet_wrap(~Species)+
  theme(strip.text = element_text(face = "italic"))

```

*H. curtus* adults had the highest mortality rate overall. *H. schistosus* juveniles had the lowest.

## Mortality rate by sex

```{r}
snakes%>%
  filter(Sex != "")%>%
  group_by(Species, Sex)%>%
  count(Condition.at.encounter..D.A.)%>%
  # Calculating mortality rate
  mutate(N = sum(n),
         prop.dead = n/N,
         sd = sqrt(prop.dead*(1-prop.dead)/N))%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

### Testing difference in mortality rates across sexes withing species

Logistic model:

```{r}

snakes%>%
  filter(Sex %in% c("Male", "Female"),
         Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Species, Sex, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")))%>%
  group_by(Species)%>%
  nest()%>%
  mutate(test = map(data, ~glm(Condition ~ Sex, family = "binomial", data = .)),
         sumr = map(test, ~broom::tidy(.)),
         stat = map(test, broom::glance),
         r2 = map(test, ~mcfadden(.)))%>%
  select(sumr, stat, r2)%>%
  unnest()%>%
  select(term:p.value, df.residual, r2)

```

Fitted probability of mortality across sexes:

```{r}
sex = data.frame(Sex = c("Male", "Female"))

snakes%>%
  filter(Sex %in% c("Male", "Female"),
         Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Species, Sex, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")))%>%
  group_by(Species)%>%
  nest()%>%
  mutate(mod = map(data, ~glm(Condition ~ Sex, family = "binomial", data = .)),
         fit = map(mod, ~predict(., sex, se.fit = T, type = "response")),
         fit = map(fit, ~data.frame(.)))%>%
  select(fit)%>%
  unnest(fit)%>%
  mutate(Sex = c("Male", "Female"))%>%
  ggplot(aes(Sex, fit))+
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit))+
  labs(y = "P(Dead)")+
  facet_wrap(~Species)+
  theme(strip.text = element_text(face = "italic"))

```

There was no difference in mortality rates across sexes in either species.

## Mortality and female reproductive status

Are gravid females more susceptible to bycatch mortality than the rest of our sample?

```{r}
HS%>%
  filter(Sex == "Female",
         Class == "Adult")%>%
  group_by(Gravid)%>%
  count(Condition.at.encounter..D.A.)%>%
  # Calculating mortality rate
  mutate(N = sum(n),
         prop.dead = n/N, 
         sd = sqrt(prop.dead*(1-prop.dead)/N))%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

### Testing thing difference in mortality rate of *H. schistosus* females by reproductive state

Logistic model:

```{r}

HS%>%
  filter(Sex == "Female",
         Class == "Adult",
         Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Gravid, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")),
         Gravid = ifelse(Gravid == "", "No", Gravid))%>%
  nest()%>%
  mutate(test = map(data, ~glm(Condition ~ Gravid, family = "binomial", data = .)),
         sumr = map(test, ~broom::tidy(.)),
         stat = map(test, broom::glance),
         r2 = map(test, ~mcfadden(.)))%>%
  select(sumr, stat, r2)%>%
  unnest()%>%
  select(term:p.value, df.residual, r2)
```

Fitted probability of mortality by reproductive state:

```{r}
gravid = data.frame(Gravid = c("Yes", "No"))

HS%>%
  filter(Sex == "Female",
         Class == "Adult",
         Condition.at.encounter..D.A. %in% c("Dead", "Alive"))%>% # removing missing data
  select(Gravid, Condition.at.encounter..D.A.)%>%
  mutate(Condition = factor(Condition.at.encounter..D.A., levels = c("Alive", "Dead")),
         Gravid = ifelse(Gravid == "", "No", Gravid))%>%
  nest()%>%
  mutate(mod = map(data, ~glm(Condition ~ Gravid, family = "binomial", data = .)),
         fit = map(mod, ~predict(., gravid, se.fit = T, type = "response")),
         fit = map(fit, ~data.frame(.)))%>%
  select(fit)%>%
  unnest(fit)%>%
  mutate(State = c("Gravid", "Not Gravid"))%>%
  ggplot(aes(State, fit))+
  geom_pointrange(aes(ymin = fit - se.fit, ymax = fit + se.fit))+
  labs(y = "P(Dead)")+
  theme(strip.text = element_text(face = "italic"))

```

# Observed reproductive cycle of *H. schistosus*

## Proportion of gravid females in the sample

```{r}
# percentage of gravid females in sample

HS%>%
  count(Gravid)%>%
  mutate(N = sum(n))%>%
  mutate(prop.gravid = n/N,
         sd = sqrt(prop.gravid*(1-prop.gravid)))%>%
  filter(Gravid == "Yes")%>%
  select(-c(Gravid))
```

## Number of gravid females encountered per year

```{r}
# checking the number of gravid females per year

HS%>%
  group_by(Year)%>%
  filter(Gravid == "Yes")%>%
  count(Gravid)%>%
  spread(Gravid, n)%>%
  rename(`n(Gravid Females)` = Yes)
```

## Monthly variation in proportion of gravid females

Proper sampling was only conducted in 2018/19 and hence only this period is used for analysis of reproductive cycles.

```{r}
# calculating the proportion of gravid females per month

gravid <- HS%>%
  mutate(Month = factor(Month, levels = month.name))%>%
  filter(Year == "2018" | Year == "2019")%>% # only for 2018/19
  group_by(Month)%>%
  summarise(N = n(),
            gravid = sum(Gravid == "Yes"),
            prop.gravid = gravid/N,
         sd = sqrt(prop.gravid*(1-prop.gravid)/N))

gravid
```

Describing change in the relative proportions of gravid females across months and years of sampling.

```{r}
# plotting prop gravid per month

gravid%>%
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(Month, fill = list(prop.gravid = 0))%>%
  ggplot(aes(Month, prop.gravid))+
  geom_pointrange(size = 1, aes(ymin = prop.gravid + sd, ymax = prop.gravid - sd))+
  geom_path(aes(group = 1), size = 1, linetype = "dashed")+
  geom_segment(aes(x = "April", xend = "April", y = 0 , yend = 0.02), #marking births
               arrow = arrow(length = unit(0.25, "cm"), ends = "first"), size = 1)+
  geom_text(aes(x = "April", y = 0.04, label = paste("Observed \n births")))+
  geom_vline(aes(xintercept = "June"), size = 1)+#start of the monsoon ban
  geom_vline(aes(xintercept = "August"), size = 1)+#end of the monsoon ban
  geom_text(aes(x = "July", y = 0.20, label = "Monsoon ban"))+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(y = "Proportion of gravid females")

```

Pregnancy from Novemeber to May. Peak in Feb. Observed two live births in April.

# Development of embryos and eggs

## Sample size

```{r}
embryos%>%
  summarise(n.mothers = length(unique(Field.Code)),
            n.embryos = length(unique(Embryo.Code)))
```

## Summary statistics of egg measurements

```{r}
embryos%>%
  select(Egg.Length..mm.., Egg.Width..mm.., Egg.Weigth..g.., Snout.to.Vent..cm., Embryo.Weight..g.)%>%
  skimr::skim()%>%
  skimr::yank("numeric")%>%
  select(c(skim_variable, mean, sd))
```

## Sex ratios within clutches

```{r}
embryos%>%
  filter(Sex != "")%>%
  summarise(prop.female = sum(Sex == "Female")/n(),
            sd = sqrt(prop.female*(1-prop.female)/n()))
```

### Number of feamle embryos

```{r}
embryos%>%
  filter(Sex != "")%>%
  summarise(females = sum(Sex == "Female"),
            N = n())
```

### Testing clutch sex ratio

```{r}
broom::tidy(prop.test(86, 166, p = 0.5))%>% # Z - test
  select(estimate:conf.high)
```

The sex ratio in clutches is not significantly different from 0.5.

## Infertility rate

Percentage of eggs with no embryos present:

```{r}
embryos%>%
  count(Embryo)%>%
  mutate(N = sum(n), rate = n/N, sd = sqrt(rate*(1-rate)/N))%>%
  select(-N)
```

## Matrotrophy

Do female *H. schistosus* expend energy in the development of embryos? or Are the eggs formed and yolk only provides nourishment?

### Change in yolk weight with embryo weight

```{r}
embryos%>%
  ggplot(aes(Embryo.Weight..g., Yolk.Weight..g.))+
  geom_point(size = 3)+
  geom_smooth(method = "lm", linetype = "dashed", size = 1)+
  labs(x = "Embryo weight (g)", y = "Yolk weight (g)")
```

**Linear model to test change in yolk weight with embryo development:**

```{r}
embryos%>%
  select(Yolk.Weight..g., Embryo.Weight..g.)%>%
  nest()%>%
  # Yolk weight is normally distributed
  mutate(mod = map(data, ~lm(Yolk.Weight..g. ~ Embryo.Weight..g., data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:r.squared, df.residual)
```

Yolk weight decreases by 0.45 g for every 1g increase in embryo weight.

### Change in egg mass with embryo development

```{r fig.height=6}
embryos%>%
  ggplot(aes(Embryo.Weight..g., Egg.Weigth..g..))+
  geom_point(aes(col = Yolk.Weight..g.), size = 3)+
  geom_smooth(method = "lm", linetype = "dashed", size = 1)+
  scale_color_viridis(name = "Yolk Weight (g)")+
  labs(x = "Embryo Weight (g)", y = "Total Egg Weight (g)")

```

Linear model to test relation ship between total egg mass and embryo development:

```{r}
embryos%>%
  select(Egg.Weigth..g.., Embryo.Weight..g.)%>%
  nest()%>%
  # Total egg mass is normally distributed
  mutate(mod = map(data, ~lm(Egg.Weigth..g.. ~ Embryo.Weight..g., data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:r.squared, df.residual)
```

Total egg mass increases by 0.36 g for every 1g increase in embryo weight.

Note: Embryo weight is log-normally distributed.

Along with decrease of fat bodies of gravid females (Figure A1) as embryos develop indicates matrotrophic nutrition.

# Reproductive effort

Does the amount of enery expended by female *H. schistosus* reduce with age?

```{r}
# Cleaning reproductive effort data

re <- embryos%>%
  left_join(snakes, by = c("Date", "Field.Code", "Species"))%>%
  select(Date, Field.Code, Embryo.Code, Species, Egg.Length..mm..:Tail.Length..cm..x,
         Egg.Weigth..g..:Sex.x, Snout.to.Vent..cm..y:Tail.Length..cm..y, Weight..g.,
         -Body.Length..cm..x, - Body.Length..cm..y)%>%
  filter(Species == "Hydrophis schistosus")%>%
  rename(Embryo.SVL = Snout.to.Vent..cm..x,
         Embryo.TL = Tail.Length..cm..x,
         Embryo.Sex = Sex.x,
         Female.SVL = Snout.to.Vent..cm..y,
         Female.TL = Tail.Length..cm..y,
         Mother.Wt = Weight..g.
         )%>%
  # Removing erroneous data point
  filter(Female.SVL > 50)
```

## Increase in clutch size with female age

```{r}
re_clutch <- re%>%
  select(Field.Code, Egg.Weigth..g.., Mother.Wt, Female.SVL, Embryo.SVL)%>%
  group_by(Field.Code)%>%
  summarise(clutch.size = n(),
            Clutch.wt = sum(Egg.Weigth..g.., na.rm = T),
            Mother.wt = last(Mother.Wt),
            Female.SVL = last(Female.SVL),
            Embryo.SVL = median(Embryo.SVL, na.rm = T))%>%
  mutate(rcm = Clutch.wt/(Mother.wt - Clutch.wt))%>%
  filter(rcm > 0)%>% # removing missing data
  replace_na(replace = list(Embryo.SVL = 3)) # filling small embryos svl

re_clutch%>%
  ggplot(aes(Female.SVL, clutch.size))+
  geom_point(size =3)+
  geom_smooth(method = 'glm', method.args = list(family = 'poisson'))+
  labs(x = "Female SVL (cm)", y = "Clutch Size")
```

**Generalised linear model to test change in clutch size with female SVL:**

Clutch size is poisson distributed with mean `r mean(re_clutch$clutch.size)` and variance `r var(re_clutch$clutch.size)`

```{r}

re_clutch%>%
  select(Female.SVL, clutch.size)%>%
  nest()%>%
  mutate(mod = map(data, ~glm(clutch.size ~ Female.SVL, data = ., family = "poisson")),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance),
         r.squared =  map_dbl(stat, ~mcfadden(.)))%>%
  select(sumr, stat,r.squared)%>%
  unnest()%>%
  select(c(term:p.value, r.squared, df.residual))

```

The number of eggs borne by females increase by `r exp(0.025)` for every 1 cm increase in female SVL.

## Change in overall reproductive effort with age

**Beta regression to determine relationship between relative clutch mass and female SVL:**

```{r}
require(betareg)

re_clutch%>%
  select(Female.SVL, Embryo.SVL, rcm)%>%
  nest()%>%
  mutate(mod = map(data, ~betareg(rcm ~ Female.SVL + Embryo.SVL, data = .)),
         sumr = map(mod, ~tidy(.)),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(component:pseudo.r.squared, df.residual)

```

```{r fig.height=6}
# Predicting for plotting fit
require(gam)
require(mgcv)

re_mod <- gam(rcm ~ Female.SVL + Embryo.SVL, family = betar(link = "logit"), data = re_clutch)

re_simdata <- data.frame(Female.SVL = seq(min(re_clutch$Female.SVL), max(re_clutch$Female.SVL), 0.1),
                         Embryo.SVL = median(re$Embryo.SVL, na.rm = T))

re_pred <- predict(re_mod, newdata = re_simdata, "response", se.fit = T)

re_simdata <- enframe(re_pred)%>%
  spread(name, value)%>%
  unnest()%>%
  mutate(Female.SVL = re_simdata$Female.SVL,
         Embryo.SVL = re_simdata$Embryo.SVL)

# Plotting

re_clutch%>%
  ggplot(aes(Female.SVL, rcm))+
  geom_point(aes(col = clutch.size), size = 3)+
  geom_smooth(data = re_simdata, aes(y = fit, ymin = fit - se.fit, ymax = fit + se.fit), stat = "identity", linetype = "dashed")+
  scale_y_continuous(name = "Relative clutch mass")+
  scale_x_continuous(limits = c(85, 115), name = "Female SVL (cm)")+
  scale_color_viridis(name = "Clutch size")+
  theme(legend.text = element_text(size = 10))
```

The overall reproductive effort **does not change** with female age.

## Change in reproductive effort per embryo with female age

Does the effort per embryo change with female age?

```{r }
# calculating reproductive effort per embruo

re_embryo <- re%>%
  select(Field.Code, Embryo.Code, Egg.Weigth..g..,  Mother.Wt, Female.SVL, Embryo.Sex, Embryo.SVL)%>%
  group_by(Field.Code)%>%
  mutate(clutch.wt = sum(Egg.Weigth..g..), 
         Female.wt = Mother.Wt - clutch.wt)%>%
  group_by(Field.Code, Embryo.Code)%>%
  summarise(inv = Egg.Weigth..g../Female.wt,# investment per embryo
            Female.SVL = last(Female.SVL),
            Embryo.Sex = Embryo.Sex,
            Embryo.SVL = Embryo.SVL)%>%
  ungroup()

```

**Linear model to determine relationship between female SVL and investment per embryo:**

We have controlled for the effect of embryo development.

```{r}
re_embryo%>%
  select(Female.SVL, inv, Embryo.SVL)%>%
  nest()%>%
  mutate(mod = map(data, ~betareg(inv ~ Female.SVL + Embryo.SVL, data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr,stat)%>%
  unnest()%>%
  select(component:p.value, pseudo.r.squared, df.residual)
```

Plotting relatonship between residuals and female SVL:

```{r fig.height=6}
# using residuals to control for embryo development

emsvlinv <- gam(inv ~ Female.SVL + Embryo.SVL, family = betar(link = "logit"), data = re_embryo)

inv_simdata <- data.frame(Female.SVL = seq(min(re_embryo$Female.SVL), max(re_embryo$Female.SVL), 0.1),
                          Embryo.SVL = median(re_embryo$Embryo.SVL, na.rm = T))%>%
  nest()%>%
  mutate(sim = map(data, ~predict(object = emsvlinv, newdata = ., type = "response", se.fit = T)),
         sim = map(sim, ~data.frame(.)))%>%
  unnest()

# Plotting

re_embryo%>%
  ggplot(aes(Female.SVL, inv))+
  geom_point(aes(col = Embryo.SVL), size = 3)+
  geom_smooth(data = inv_simdata, aes(y = fit, ymin = fit -se.fit, ymax = fit+se.fit),stat = "identity", linetype = "dashed", size = 1)+
  scale_y_continuous(name = "Relative egg mass")+
  scale_x_continuous(limits = c(85, 115), name = "Female SVL (cm)")+
  scale_color_viridis(name = "Embryo SVL (cm)")


```

The relative egg mass (controlled for embryo development) reduces with female SVL.

## Difference in reproductive effort for male and female embryos

Does female investment differ by sex of the embryo?

**Linear model to test for the difference in female investment by embryo sex:**

We controlled for the effect of embryo development.

```{r}
re_embryo%>%
  filter(Embryo.Sex == "Male" | Embryo.Sex == "Female")%>%
  select(Embryo.Sex, inv, Embryo.SVL)%>%
  nest()%>%
  mutate(mod = map(data, ~betareg(inv ~ Embryo.Sex + Embryo.SVL, data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(component:p.value, pseudo.r.squared, df.residual)
```

```{r fig.height=6}
# Simulating data to plot model fit

inv_sex_mod <- gam(inv ~ Embryo.Sex + Embryo.SVL, family = betar(link = "logit"), data = re_embryo)
  
inv_sex_simdata <- data.frame(Embryo.Sex = c("Male", "Female"),
                              Embryo.SVL = median(re_embryo$Embryo.SVL, na.rm = T))%>%
  nest()%>%
  mutate(sim = map(data, ~predict(inv_sex_mod, newdata = ., type = "response", se.fit = T)),
         sim = map(data, ~data.frame(sim)))%>%
  unnest()

re_embryo%>%
  filter(Embryo.Sex == "Male" | Embryo.Sex == "Female")%>%
  ggplot(aes(Embryo.Sex, inv))+
  geom_jitter(aes(col = Embryo.SVL), size = 2, width = 0.125)+
  geom_pointrange(data = inv_sex_simdata, aes(y = fit, ymin = fit -(se.fit*1.96), ymax = fit+(se.fit*1.96)), size = 1, col = "red")+
  scale_y_continuous(name = "Relative egg mass")+
  scale_x_discrete(name = "Embryo Sex")+
  scale_color_viridis(name = "Embryo SVL (cm)")

```

Female investment does not differ significantly with embryo sex.
