---
title: "Demographics and reproductive biology of H. schistosus from bycatch on the west coast of India"
subtitle: "Analysis for Manuscript"
output:
   html_document:
    css: style.css
    df_print: kable
    theme: spacelab
    toc: yes
    toc_float: yes
    toc_depth: 2
    number_sections: yes
---

```{r setup, include=FALSE}
#Global options

options(max.print="75")
knitr::opts_chunk$set(echo = FALSE, error = T, warning = F, message = F, tidy = T, cache = T,
                      dpi = 300, fig.align = "center", fig.height = 4.5, fig.width = 8)

#Required libraries

library(tidyverse)
library(lubridate)
library(viridis)

# Importing data

snakes = read.csv("./Data/Sea-snakes_fish-dep_mastersheet_250420.csv")
embryos = read.csv("./Data/Sea snake_embryo data_mastersheet.csv")

snakes <- snakes%>%
  mutate(Year = as.factor(Year))%>%
  filter(Species == "Hydrophis curtus" | Species == "Hydrophis schistosus")%>%
  droplevels()


theme_set(theme_bw()+
            theme(legend.position = "top",
                  text = element_text(size = 15)
                  )
          )#setting ggplot theme

HS <- snakes%>%
  filter(Species == "Hydrophis schistosus")%>% # filtering out other snakes from bycatch data
  mutate(Year = as.factor(Year))


```

# Demographics of _H. schistosus_ across time

## Sample size

```{r}
snakes%>%
  group_by(Species)%>%
  count(name = "N")
```

```{r}
snakes%>%
  filter(Gear.Type != "")%>%
  select(Date, Gear.Type, No..of.Hauls, Average.Haul.Duration..Hours., Tow.duration.hours)%>%
  distinct()%>%
  mutate(Tow.duration.hours = case_when(!is.na(Tow.duration.hours) ~ Tow.duration.hours,
                                        is.na(Tow.duration.hours) ~ No..of.Hauls*Average.Haul.Duration..Hours.))%>%
  group_by(Gear.Type)%>%
  summarise(n = n(),
            haul.hours = sum(Tow.duration.hours, na.rm = T))
  
```

## Age structure

```{r}
age.yr <- snakes%>%
  group_by(Species, Year)%>%
  summarise(N = n(),
            mean = mean(Snout.to.Vent..cm., na.rm = T))


maturtity <- snakes%>%
  group_by(Species)%>%
  count()%>%
  mutate(juv = 35,
         adult = ifelse(Species == "Hydrophis curtus", 54, 65))
  
```

```{r}
# plotting distribution of SVL across years

snakes%>%
  filter(Snout.to.Vent..cm. > 20)%>%
  ggplot(aes(Year, Snout.to.Vent..cm.))+
  geom_violin(fill = "grey")+
  geom_boxplot(width = 0.1)+
  geom_hline(data = maturtity, aes(yintercept = adult), linetype = "dashed")+
  geom_hline(data = maturtity, aes(yintercept = juv), linetype = "dotted")+
  #stat_summary(fun.data = "mean_sdl", geom = "pointrange", size = 1)+
  geom_label(data = age.yr, aes(Year, 10, label = N))+
  facet_wrap(~Species, ncol = 1, scales = "free_y")+
  labs(y = "Snout to vent length (cm)")+
  theme(strip.text = element_text(face = "italic"))

ggsave(last_plot(), filename = "./Figures/figure1.tiff", height = 6, width = 8)
  
```

```{r}
pop.yr <- snakes%>%
  filter(!is.na(Class))%>% # fix 2016 svl
  group_by(Species, Year)%>%
  count(Class)%>%
  complete(Class, fill = list(n = 0))%>%
  mutate(N = sum(n))%>%
  group_by(Year, Species, Class)%>%
  mutate(prop.age = n/N)

pop.yr%>%
  group_by(Species, Class)%>%
  skimr::skim(prop.age)%>%
  skimr::yank("numeric")%>%
  select(-hist)
```

```{r}
pop.yr%>%
  filter(Year != 2019)%>%
  group_by(Species, Class)%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$n, .$N)),
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()%>%
  select(Class:p.value)
```


Age structure does not change significantly over a four year period from 2016 to 2019.

```{r}

library(car)

snakes%>%
  filter(Year != "2019")%>%
  group_by(Species)%>%
  select(Year, Snout.to.Vent..cm.)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(Snout.to.Vent..cm. ~ Year, data = .)),
          sumr = map(mod, broom::tidy),
          stat = map(mod, car::Anova))%>%
  select(stat)%>%
  unnest()
```

GLM shows significance but how do you interpret these results?

```{r, fig.height = 4.5, fig.width=8}
# plotting distribution of SVL across months

month.svl <- HS%>%
  filter(Year == "2018", # data not sufficient for other years
         Snout.to.Vent..cm. > 20)%>% # removing erroneous data
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(Month)%>%
  group_by(Month)%>%
  summarise(mean.SVL = mean(Snout.to.Vent..cm., na.rm = T))

births <- data.frame(Species = "Hydrophis schistosus", Month = "April")
  
HS%>%
  filter(Year == "2018", # data not sufficient for other years
         Snout.to.Vent..cm. > 20)%>% # removing erroneous data
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(nesting(Species), Month)%>%
  droplevels()%>%
  ggplot(aes(Month, Snout.to.Vent..cm.))+
  geom_violin(fill = "light grey")+
  #geom_point(data = month.svl, aes(x = Month, y = mean.SVL), size = 3)+
  geom_boxplot(width = 0.1)+
  geom_segment(data = births, aes(x = Month, xend = Month, y = 0 , yend = 10), #marking births
               arrow = arrow(length = unit(0.25, "cm"), ends = "first"), size = 1)+
  geom_text(data = births, aes(x = Month, y = 20, label = paste("Observed \n births")))+
  geom_vline(aes(xintercept = "June"), size = 1)+#start of the monsoon ban
  geom_vline(aes(xintercept = "August"), size = 1)+#end of the monsoon ban
  geom_text(aes(x = "July", y = 80, label = "Monsoon ban"))+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(y = "Snout to vent length (cm)")

```

Proportion of neonates increases in March to May. Birth and neonates were observed around the same time.

- GLM testing change in proportion of SVL < 40 over months
- Other ways to test age structure?

```{r}
# proportion of jeuveniles in each month

snakes%>%
  filter(Year == "2018", # data not sufficient for other years
         Snout.to.Vent..cm. > 20)%>% # removing erroneous data
  mutate(Month = factor(Month, levels = month.name))%>%
  group_by(Species, Month)%>%
  summarise(prop.neonate = sum(Snout.to.Vent..cm. < 40)/n())%>%
  spread(Month, prop.neonate)
```


## Sex ratios

```{r, fig.height=8}
# Plotting sex ratios across years

snakes%>%
  filter(Sex == "Male" | Sex == "Female")%>%
  group_by( Species, Year, Month)%>%
  summarise(N = n(),
            females = sum(Sex == "Female"),
            prop.female = females/N)%>%
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(Month)%>%
  ggplot(aes(Month, prop.female))+
  geom_col(width = 0.5, col = "black")+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  facet_grid(Year ~ Species)

```

Proportion of females to entire population stays constant over the sampling period.

```{r}
snakes%>%
  group_by(Species)%>%
  filter(Sex == "Male" | Sex == "Female")%>%
  summarise(females = sum(Sex == "Female"),
            N = n())

# is proportion of females different from 0.5?
snakes%>%
  filter(Sex == "Male" | Sex == "Female")%>%
  group_by(Species)%>%
  summarise(females = sum(Sex == "Female"),
            N = n())%>%
  group_by(Species)%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$females, .$N, p = 0.5)),
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()
```

Sex ratio is not different from 0.5 p = 0.71

# Observed reproductive cycle of _H. schistosus_

```{r}
# percentage of gravid females in sample

HS%>%
  count(Gravid)%>%
  mutate(N = sum(n))%>%
  mutate(prop.gravid = n/N)%>%
  filter(Gravid == "Yes")
```

̥
```{r}
# checking the number of gravid females per year

HS%>%
  group_by(Year)%>%
  filter(Gravid == "Yes")%>%
  count(Gravid)%>%
  spread(Gravid, n)
```

Proper sampling was only conducted in 2018/19 and hence only this period is used for analysis of reproductive cycles.

```{r}
# calculating the proportion of gravid females per month

gravid <- HS%>%
  mutate(Month = factor(Month, levels = month.name))%>%
  filter(Year == "2018" | Year == "2019")%>% # only for 2018/19
  group_by(Month)%>%
  summarise(N = n(),
            gravid = sum(Gravid == "Yes"),
            prop.gravid = gravid/N)

print(gravid)
```

Describing change in the relative proportions of gravid females across months and years of sampling.

```{r}
# plotting prop gravid per month

gravid%>%
  mutate(Month = factor(Month, levels = month.name))%>%
  complete(Month, fill = list(prop.gravid = 0))%>%
  ggplot(aes(Month, prop.gravid))+
  geom_point(size = 3)+
  geom_path(aes(group = 1), size = 1, linetype = "dashed")+
  geom_segment(aes(x = "April", xend = "April", y = 0 , yend = 0.02), #marking births
               arrow = arrow(length = unit(0.25, "cm"), ends = "first"), size = 1)+
  geom_text(aes(x = "April", y = 0.04, label = paste("Observed \n births")))+
  geom_vline(aes(xintercept = "June"), size = 1)+#start of the monsoon ban
  geom_vline(aes(xintercept = "August"), size = 1)+#end of the monsoon ban
  geom_text(aes(x = "July", y = 0.20, label = "Monsoon ban"))+
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  labs(y = "Proportion of gravid females")

```

Pregnancy from Novemeber to May. Peak in Feb. Observed two live births in April.

- Anova/GLM with change in proportion of gravid females over time.

# Development of embryos and eggs

## Sample size

```{r}
embryos%>%
  summarise(n.mothers = length(unique(Field.Code)),
            n.embryos = length(unique(Embryo.Code)))
```

## Summary

```{r}
embryos%>%
  select(Egg.Length..mm., Egg.Width..mm., Egg.Weigth..g., Snout.to.Vent..cm., Embryo.Weight..g.)%>%
  skimr::skim()%>%
  skimr::yank("numeric")%>%
  select(-hist)
```

```{r}
embryos%>%
  group_by(Field.Code)%>%
  filter(Sex != "")%>%
  summarise(prop.female = sum(Sex == "Female")/n())%>%
  skimr::skim(prop.female)%>%
  skimr::yank("numeric")%>%
  select(-hist)

embryos%>%
  filter(Sex != "")%>%
  summarise(females = sum(Sex == "Female"),
            N = n())

broom::tidy(prop.test(86, 166, p = 0.5))%>%
  select(estimate:conf.high)
```

The sex ratio in clutches is not significantly different from 0.5.

## Infertility 

```{r}
embryos%>%
  count(Embryo)%>%
  mutate(N = sum(n), rate = n*100/N)
```

## Matrotrophy

Do female _H. schistosus_ expend energy in the development of embryos? or Are the eggs formed and yolk only provides nourishment?

```{r}
embryos%>%
  ggplot(aes(Embryo.Weight..g., Yolk.Weight..g.))+
  geom_point(size = 3)+
  geom_smooth(method = "lm", linetype = "dashed", size = 1)
```

```{r}
embryos%>%
  select(Yolk.Weight..g., Embryo.Weight..g.)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(Yolk.Weight..g. ~ Embryo.Weight..g., data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:r.squared)
```

Yolk weight decreases as embryo weight increases. 

```{r}
embryos%>%
  ggplot(aes(Embryo.Weight..g., Egg.Weigth..g.))+
  geom_point(aes(col = Yolk.Weight..g.), size = 3)+
  geom_smooth(method = "lm", linetype = "dashed", size = 1)+
  scale_color_viridis(name = "Yolk Weight (g)")+
  labs(x = "Embryo Weight (g)", y = "Total Egg Weight (g)")

```

```{r}
embryos%>%
  select(Egg.Weigth..g., Embryo.Weight..g.)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(Egg.Weigth..g. ~ Embryo.Weight..g., data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:r.squared)
```

Egg weight increases as embryo weight increases

Decrease of fat bodies as embryos develop indicates matrotrophic nutrition.

- Photos of females dissected ordered according to embryo dev stage 

# Reproductive effort

```{r}
# Cleaning reproductive effort data

re <- embryos%>%
  left_join(snakes, by = c("Date", "Field.Code", "Species"))%>%
  select(Date, Field.Code, Embryo.Code, Species, Egg.Length..mm.:Tail.Length..cm..x,
         Egg.Weigth..g.:Sex.x, Snout.to.Vent..cm..y:Tail.Length..cm..y, Weight..g.,
         -Body.Length..cm..x, - Body.Length..cm..y)%>%
  filter(Species == "Hydrophis schistosus")%>%
  rename(Embryo.SVL = Snout.to.Vent..cm..x,
         Embryo.TL = Tail.Length..cm..x,
         Embryo.Sex = Sex.x,
         Female.SVL = Snout.to.Vent..cm..y,
         Female.TL = Tail.Length..cm..y,
         Mother.Wt = Weight..g.
         )%>%
  filter(Female.SVL > 50)
```


## Increase in clutch size with female age

Does the amount of enery expended by female _H. schistosus_ reduce with age?

```{r}
re_clutch <- re%>%
  select(Field.Code, Egg.Weigth..g., Mother.Wt, Female.SVL)%>%
  group_by(Field.Code)%>%
  summarise(clutch.size = n(),
            Clutch.wt = sum(Egg.Weigth..g.),
            Mother.wt = last(Mother.Wt),
            Female.SVL = last(Female.SVL))%>%
  mutate(rcm = Clutch.wt/(Mother.wt - Clutch.wt))
```

```{r}
re_clutch%>%
  ggplot(aes(Female.SVL, clutch.size))+
  geom_point(size =3)+
  geom_smooth(method = "lm")
```

```{r}
re_clutch%>%
  select(Female.SVL, clutch.size)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(clutch.size ~ Female.SVL, data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:r.squared)
```

The number of eggs borne by females increases with age (SVL).

## Change in overall reproductive effort with age

```{r}
re_clutch%>%
  ggplot(aes(Female.SVL, rcm))+
  geom_point(aes(col = clutch.size), size = 3)+
  geom_smooth(method = "lm", linetype = "dashed")+
  scale_y_continuous(name = "Relative clutch mass")+
  scale_x_continuous(limits = c(85, 115), name = "Female SVL (cm)")+
  scale_color_viridis(name = "Clutch size")+
  theme(legend.text = element_text(size = 10))
```

```{r}
re_clutch%>%
  select(Female.SVL, rcm)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(rcm ~ Female.SVL, data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:r.squared)
```

The overall reproductive effort **does not change** with female age.

## Change in reproductive effort per embryo with female age

Does the effort per embryo change with female age?

```{r}
# reproductive effort per embruo

re_embryo <- re%>%
  select(Field.Code, Embryo.Code, Egg.Weigth..g.,  Mother.Wt, Female.SVL, Embryo.Sex, Embryo.SVL)%>%
  group_by(Field.Code)%>%
  mutate(clutch.wt = sum(Egg.Weigth..g.), 
         Female.wt = Mother.Wt - clutch.wt)%>%
  group_by(Field.Code, Embryo.Code)%>%
  summarise(inv = Egg.Weigth..g./Female.wt, 
            Female.SVL = last(Female.SVL),
            Embryo.Sex = Embryo.Sex,
            Embryo.SVL = Embryo.SVL)%>%
  ungroup()
```

```{r}
# using residuals to control for embryo development

emsvlinv <- lm(inv ~ Embryo.SVL, data = re_embryo)

re_embryo%>%
  modelr::add_residuals(emsvlinv)%>%
  ggplot(aes(Female.SVL, resid))+
  geom_point(aes(col = Embryo.SVL), size = 3)+
  geom_smooth(method = "lm", linetype = "dashed", size = 1)+
  scale_y_continuous(name = "Relative egg mass (residuals)")+
  scale_x_continuous(limits = c(85, 115), name = "Female SVL (cm)")+
  scale_color_viridis(name = "Embryo SVL (cm)")

```

```{r}
re_embryo%>%
  select(Female.SVL, inv, Embryo.SVL)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(inv ~ Female.SVL + Embryo.SVL, data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:p.value, r.squared)
```

The relative egg mass (controlled for embryo development) reduces with female SVL.

## Difference in reproductive effort for male and female embryos

```{r}
re_embryo%>%
  filter(Embryo.Sex == "Male" | Embryo.Sex == "Female")%>%
  modelr::add_residuals(emsvlinv)%>%
  ggplot(aes(Female.SVL, resid, shape = Embryo.Sex))+
  geom_point(aes(col = Embryo.SVL), size = 3)+
  geom_smooth(method = "lm", linetype = "dashed", size = 1)+
  scale_y_continuous(name = "Relative egg mass (residuals)")+
  scale_x_continuous(limits = c(85, 115), name = "Female SVL (cm)")+
  scale_color_viridis(name = "Embryo SVL (cm)")
```

```{r}
re_embryo%>%
  filter(Embryo.Sex == "Male" | Embryo.Sex == "Female")%>%
  select(Embryo.Sex, inv, Embryo.SVL)%>%
  nest()%>%
  mutate(mod = map(data, ~lm(inv ~ Embryo.Sex + Embryo.SVL, data = .)),
         sumr = map(mod, broom::tidy),
         stat = map(mod, broom::glance))%>%
  select(sumr, stat)%>%
  unnest()%>%
  select(term:p.value, r.squared)
```

# Mortality in bycatch

## Overall mortality

```{r}
snakes%>%
  group_by(Species)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

```{r}
snakes%>%
  group_by(Species)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)%>%
  ungroup()%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$n, .$N)),
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()%>%
  select(estimate1:p.value)
```

## Mortality by age class

```{r}
snakes%>%
  filter(!is.na(Class))%>%
  group_by(Species, Class)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

```{r}
snakes%>%
  filter(!is.na(Class))%>%
  group_by(Species, Class)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)%>%
  group_by(Species)%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$n, .$N)),
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()%>%
  select(estimate1:p.value)

```

## Mortality by sex

```{r}
snakes%>%
  filter(Sex != "")%>%
  group_by(Species, Sex)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

```{r}
snakes%>%
  filter(Sex != "")%>%
  group_by(Species, Sex)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)%>%
  group_by(Species)%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$n, .$N)),
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()%>%
  select(estimate1:p.value)
```

## Mortality by female reproductive status

```{r}
HS%>%
  filter(Sex == "Female",
         Class == "Adult")%>%
  group_by(Gravid)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)
```

```{r}
HS%>%
  filter(Sex == "Female",
         Class == "Adult")%>%
  group_by(Gravid)%>%
  count(Condition.at.encounter..D.A.)%>%
  mutate(N = sum(n),
         prop.dead = n/N)%>%
  filter(Condition.at.encounter..D.A. == "Dead")%>%
  select(-Condition.at.encounter..D.A.)%>%
  ungroup()%>%
  nest()%>%
  mutate(test = map(data, ~prop.test(.$n, .$N)),
         sumr = map(test, broom::tidy))%>%
  select(sumr)%>%
  unnest()%>%
  select(estimate1:p.value)
```

